Comparison of MCMC algorithms in Stochastic Volatility Models

### 2023-10-04 manuscript/main.tex
Rank hist examples

        \subsubsection{Rank statistics and chi squared statistics}
            Rank statistics as described in the SBC section are used to evaluate the calibration of a posterior. Histograms will be used to evaluate the distribution of rank statistics. If a posterior is well calibrated then it is expected that the histogram is uniform.

            A drawback of this approach is there are more parameters than data points in this model. An alternative to visually checking for uniformity of all the parameters is to calculate the chi squared statistics for the counts in each histogram bin. Let $b_j$ be the number of counts and $e_j$ the expected count in bin $j$. Then the chi squared statistic is given by:

>>>>

        \subsubsection{Rank statistics}
            As discussed in the SBC section, rank statistics are used to evaluate the calibration of the MCMC algorithm. If a posterior is well calibrated then it is expected that the distribution of rank statistics is uniform.

            The shape of the rank statistic distribution gives insight into how a MCMC may be miscalibrated. Specifically, it gives information about the type of bias in the posterior estimates for a given parameter. Figures \ref{fig:overestimate}, \ref{fig:underestimation} and \ref{fig:underdispersed} gives some examples of the posterior estimates which result in bias of the rank statistics.

            Figure \ref{fig:overestimate} shows a distribution of rank statistics with a large left peak. This is consistent with posterior samples overestimating the true parameter. The sum of indicator random variables is small since the majority of posterior draws are larger than the true value. This results in a small rank statistic. Hence, the shape of the histogram is consistent with multiple posterior estimates as part of the SBC procedure being biased towards overestimation of the true parameter.

            Figure \ref{fig:underestimation} displays the converse behaviour. The large right peak is consistent with a bias towards under estimation of the true parameter.

            Figure \ref{fig:underdispersed} has large peaks on both ends of the histogram. The sampler is disproportionately over and under estimating the true parameter resulting in biased rank statistics. \citet{talts2018validating} describe this bias as underdispersion of the posterior relative to the prior distribuion (since the posterior draws should follow the same distribution as the prior in this simulation design). The estimated posterior is too narrow relative to the spread of the prior resulting in bias on both ends of the rank statistics distribution.

            \begin{figure}[H]
                \centering
                \includegraphics[scale=0.07]{methodology/lhs.png}
                \caption{Left: Non uniform rank statistics with peak on the left side. Right: Posterior distribution vertical line representing true parameter. Bias in rank statistics distribution comes from disproportionate number of SBC iterations overestimating the true parameter.}
                \label{fig:overestimate}
            \end{figure}

            \begin{figure}[H]
                \centering
                \includegraphics[scale=0.07]{methodology/rhs.png}
                \caption{Left: Non uniform rank statistics with peak on the right side. Right: Posterior distribution vertical line representing true parameter. Bias in rank statistics distribution comes from disproportionate number of SBC iterations underestimating the true parameter.}
                \label{fig:underestimation}
            \end{figure}

            \begin{figure}[H]
                \centering
                \includegraphics[scale=0.07]{methodology/underdispersed.png}
                \caption{Left: Non uniform rank statistics with peaks on both ends. Right: Overlayed posterior and prior distributions for the same arbitrary parameter (taken from \citet{talts2018validating}). Bias in ranks comes from underdispersed posterior distribution relative to prior.}
                \label{fig:underdispersed}
            \end{figure}

            \subsubsection{Chi squared statistics}
            A drawback from visualising the rank statistic distributions is there are more parameters (1003 parameters) than data points in this model. An alternative to visually checking for uniformity of all the parameters is to calculate the chi squared statistics for the counts in each histogram bin. Let $b_j$ be the number of counts and $e_j$ the expected count in bin $j$. Then the chi squared statistic is given by:
